Internet of Things: a Real-World Deployment Towards Smart Grids Integration

### 2013-06-28 doc/thesis/report.tex

%COVER PAGE
\input{cover_page}
\chapter{} % dedications chapter
\begin{dedication}
To my dear parents for their patience, love and support \\
To my brothers Hossein, Omar and Saad \\
To my cousin Sara and her family \\
To my family \\
To my friends
\end{dedication}
\chapter{Acknowledgements}
\paragraph{}
I am deeply thankful to my supervisor Dr. Mohamed Riduan Abid for his very close supervision, guidance and for the advice he kept giving me during the lifetime of the project. Your advices whether they are computer science related, or even every day life related wise words things I will never forget.
\paragraph*{}
I am also thankful for having Dr. Driss Benhaddou as a supervisor. His advice, insight, guidance and very enlightening ideas. I wouldn't have been so interested in smart grids without the discussions we had on the topic.
\paragraph{}
I would also like to thank Dr. Michael Gerndt for his guidance in this thesis and also for his help and advising throughout my masters degree in Technische Universät München.
\paragraph{}
I  also thank Dr. Hamid Harroud who suggested several directions to take throughout the course of this project.
\paragraph{}
My special thanks to my close friends who helped relax in times that seemed very stressful. I wouldn't have made it without their help.
\paragraph{}
I also greatly thank all excellent professors of whom I had the chance to be a student. I learned things in class  I would have never learnt elsewhere.

@@

\section{Outline of the Thesis}
\paragraph{}
The thesis is is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
\paragraph{Chapter 1}
It is an introductory chapter that gives a flavor of the project that is to be studied in this thesis and also introduces the problem and the solution proposed in this topic. This chapter also puts the reader into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}. 
\paragraph{Chapter 2}
It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}. It also discussed related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
\paragraph{Chapter 3}
It is dedicated to the \glsreset{iot}\gls{iot}. It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet, the future of Internet and also how will \gls{sg} benefit from it.
\paragraph{Chapter 4}
It will present the architecture of the system and the test-bed. It will discuss the system from different aspects, and how were the two problems solved from the architectural aspect.
\paragraph{Chapter 5}
This chapter will mostly discuss the implementation system. It will start by presenting the hardware used in this project, then move to the technology enablers. Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
\paragraph{Chapter 6}
It will start first by presenting the findings then move to explain the experiment that was done to measure the performance of the system. The data gathering section will show how and what data was gathered and then move to analyzing and interpreting the data.
\paragraph{Chapter 7, 8}
These are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.
\chapter{Literature review}
\section{The need for smart grid}
\paragraph{}
The actual electrical grid is based on a technology that is at least one century old \cite{ref5}. As a result, there is a need to change as to meet to 21st century requirement such as having more green energy in order to reduce the pollution caused mainly by the use of fossil energy. As a result, research has introduced a new technology called \glsreset{sg}\gls{sg}. Although it is still a research project and numerous issues related to \gls{sg} are left unanswered, nations across the globe started to deploy \gls{sg} project. It is also noted that in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.
\paragraph{}
The smart grid brings around new concepts that will achieve efficiency, reliable and security to the power grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6}

>>>>

%COVER PAGE
\input{cover_page}
\chapter{} % dedications chapter
\begin{dedication}
To my dear parents for their patience, love and support \\
To my brothers Hossein, Omar and Saad \\
To my cousin Sara and her family \\
To my family \\
To my friends \\

\end{dedication}
\chapter{Acknowledgements}
\paragraph{}
I am deeply thankful to my supervisor Dr. Mohamed Riduan Abid for his very close supervision, guidance and for the advice he kept giving me during the lifetime of the project. Your advices whether they are computer science related, or even every day life related wise words things I will never forget.
\paragraph*{}
I am also thankful for having Dr. Driss Benhaddou as a supervisor. His advice, insight, guidance and very enlightening ideas. I wouldn't have been so interested in smart grids without the discussions we had on the topic.
\paragraph{}
I would also like to thank Dr. Michael Gerndt for his guidance in this thesis and also for his help and advising throughout my masters degree in Technische Universät München.
\paragraph{}
I  also thank Dr. Hamid Harroud who suggested several directions to take throughout the course of this project.
\paragraph{}
My special thanks to my close friends who helped relax in times that seemed very stressful. I wouldn't have made it without their help.
\paragraph{}
I also greatly thank all excellent professors of whom I had the chance to be a student. I learned things in class  I would have never learnt elsewhere.

@@

\section{Outline of the Thesis}
\paragraph{}
The thesis is is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
\paragraph{Chapter 1}
It is an introductory chapter that gives a flavor of the project that is to be studied in this thesis and also introduces the problem and the solution proposed in this topic. This chapter also puts the reader into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}. 
\paragraph{Chapter 2}
It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}. It also discussed related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
\paragraph{Chapter 3}
It is dedicated to the \glsreset{iot}\gls{iot}. It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet, the future of Internet and also how will \gls{sg} benefit from it.
\paragraph{Chapter 4}
It will present the architecture of the system and the test-bed. It will discuss the system from different aspects, and how were the two problems solved from the architectural aspect.
\paragraph{Chapter 5}
This chapter will mostly discuss the implementation system. It will start by presenting the hardware used in this project, then move to the technology enablers. Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
\paragraph{Chapter 6}
It will start first by presenting the findings then move to explain the experiment that was done to measure the performance of the system. The data gathering section will show how and what data was gathered and then move to analyzing and interpreting the data.
\paragraph{Chapter 7, 8}
These are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.
\chapter{Literature review}
\section{The need for Smart Grid}
\paragraph{}
The actual electrical grid is based on a technology that is at least one century old \cite{ref5}. As a result, there is a need to change as to meet to 21st century requirement such as having more green energy in order to reduce the pollution caused mainly by the use of fossil energy. As a result, research has introduced a new technology called \glsreset{sg}\gls{sg}. Although it is still a research project and numerous issues related to \gls{sg} are left unanswered, nations across the globe started to deploy \gls{sg} project. It is also noted that in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.
\paragraph{}
The smart grid brings around new concepts that will achieve efficiency, reliable and security to the power grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduce the losses of energy in the transmission.  This distribution by itself enables the power grid this to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer but households can now produce and even sell the extra power. This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producer and sell the excess power to the grid \cite{ref7}. This will allow the power grid to provide power to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}
\section{\gls{iot} and \gls{sg}}
\paragraph{}
The \glsreset{iot}\gls{iot} has known high acceptance by the players in the Internet. Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to read 50 billion connected devices by the end of the decade \cite{ref3}. \gls{iot} brings several benefits to the actual Internet and these benefits will pull \gls{sg} to be part of it. Consequently, research has taken this direction which is to integrate the \gls{sg} within the \gls{iot} as most aspects of \gls{sg} will be enhanced by \gls{iot} \cite{ref8}.
\paragraph{}
\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power lines, WI-FI, Zigbee and other technologies \cite{ref8}. The smart grid's architecture based on \gls{iot} will be divided into three layers:
\begin{itemize}
\item Perception layer: Devices will autonomously perform their tasks such as sensing and executing orders.
\item Communication layer: The communication infrastructure is defined here and issues related to how to communicate information is handled.
\item Application layer: Applications can be built to make use of the underlying data.
\end{itemize}
\paragraph{}
Some of the research is discussing the need to replace Zigbee by Wi-Fi as the Wi-Fi offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10} but still Zigbee has the sole advantage of very low energy consumption which is something required in \gls{sg} and \gls{iot}, because in systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices, one cannot manage to move for the sole purpose of providing energy often. 
\paragraph{}
From the architectural point of view, integrating \gls{sg} within \gls{iot} means having to address heterogeneity issues. An \gls{iot} gateway system that is based on Zigbee and GPRS protocols \cite{ref11}. This enables then to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network. Another solution to the heterogeneity problem is proposed where the solution is to present a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices \cite{ref12}. This protocol is platform-independent, low energy consumer. Some researchers claim that the heterogeneity is mainly coming from the fact that there are different \gls{wsn} devices that do not necessarily use the same standards and protocols. They propose to move all the \glspl{wsn} to all-IP networks as this would remove most of the heterogeneity but there are already deployed legacy \glspl{wsn} \cite{ref13}. They sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6}.\glsreset{sg}
\section{The \gls{sg} two-way communication}
\paragraph{}
